### 2019-07-23 article4_tommylars.rnw
initial commit

%https://www.gnu.org/software/emacs/manual/html_node/emacs/Fill-Commands.html
%https://academic.oup.com/journals/pages/authors/latex_files
%latexmk -pdf -pvc -auxdir=/Users/larshovdanmolden/Documents/Output/  article2_mancap.tex

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
\usepackage{subfig}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{filecontents}
\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

%%%% TURN OFF PAGE NUMBER FOR INCLUSION IN THESIS
%\pagenumbering{gobble}
%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Evolution of (dynamic) capabilities and its performance implications: An empirical test and extension}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
%\author{Lars Hovdan Molden \fnref{myfootnote}}
%\address{Kongensgt 42, 7713 Steinkjer}
%\fntext[myfootnote]{Nord University Business School}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
  Despite considerable interest from the research community the theory of dynamic
  capabilities is still very much evolving. Several recent literature reviews demonstrate
  the need for better understanding of the antecedents and outcome of dynamic
  capabilities. Building on a much cited theoretical framework developed by
  \cite{zollo2002} and longitudinal dataset of Norwegian technology firms, we explore
  the role of learning as an antecedent to dynamic capabilities, and relates this to the
  outcome of dynamic capabilities themselves.

\end{abstract}

\begin{keyword}
\texttt{Dynamic Capabilities, Deliberate Learning, Dynamics }
\end{keyword}

\end{frontmatter}

%\linenumbers

\doublespacing

\section{Introduction}
There is ongoing interest in (dynamic) capabilities and how they influence firm
performance \citep{Fainshmidt2016a,Teece2014}. Indeed, \emph{dynamic capabilities} (DC),
and its front runner, the resource-based view \citep{Barney2001,Helfat2007}, is now
considered one of the major theories in strategic management
\citep{Schilke2018}. Arguably, the key reason is that this theory clearly articulates how
firms can alter their operating routines through dynamic capabilities. Thus, a key
ingredient in this theory is that DCs can alter firms operating routines, with subsequent
performance implications.

An issue that has received far less attention, however, is whether dynamics capabilities
themselves undergo an evolution \citep{Schilke2018}. Moreover, if DCs undergo an
evolution, is the evolution inert and past-dependent, or more deliberate and represent
purposeful change in firms DCs over time? Unpacking this is crucial to the development of
DC theory as it goes into the very nature of the issue as to whether and to what extent
DCs can purposefully alter firms operating routines, and the performance implications
stemming from such change \citep{Helfat2007}. It is therefore surprising that the
literature has dealt with it sparingly. What is the exception to this is \cite{Zollo2002a}
highly influential model over DC evolution where they postulate that (dynamic)
capabilities evolve with regard to firms’ \emph{deliberate learning} (DL). However, Zollo
\& Winter`s insightful theorizing has received scant empirical testing. Moreover, their
model can be developed further, by incorporating insights from subsequent research and
theorizing since this model was published. One salient issue in this regard is that Zollo
\& Winter did not consider how DCs can influence firms’ DL. Moreover, their model does not
consider how firm performance may be influenced by the evolution of (dynamic)\todo{should
  we really go with the parantheses around dynamic throuhgt the whole paper - it feels
  like we are not certain about the concept} capabilities (through DL). Such feedback
loops may significantly alter how we understand the relationship between deliberate
learning and evolution of (dynamic) capabilities in extant DC theory, including the
performance implications of such evolution. Reflecting this gap in our understanding, we
ask the following research question:

\emph{What is the role of deliberate learning in the evolution of (dynamic) capabilities and the
performance implications for such evolution?}

To answer this research question, we not only empirically test ZW’s a highly influential
model of the evolution of (dynamic) capabilities, but also extend it by incorporating (a)
feedback loops between DCs and deliberate learning and (b) incorporating firm performance
into the model.

\section{Theory}\label{sec:theory}
Firm capabilities can roughly be divided into two categories. First, ordinary capabilities
(OR) “are directed toward maintaining and leveraging the status quo in terms of the scale
and scope of activities, businesses, product lines, customer segments”
\citep[p. 393]{Schilke2018}. In other words these capabilities are the ones who “permit a
firm to ‘make a living’ in the short term” \citep[p. 991]{Winter2003}. On the other hand,
dynamic capabilities (DC) are “those that operate to extend, modify or create ordinary
capabilities” (ibid). DC theory is emerging as an important paradigm in strategic
management research, but also facing some considerable challenges with respect to its
antecedents and outcomes \citep{Eriksson2014,Schilke2018}, as well as its evolution
and development trajectory \citep{Romme2010,Zollo2002a}. Extant
literature has dealt with antecedents in many forms such as ### (ref) etc. Similarly,
outcomes have been the focal point of several contributions focusing on ### (ref)
etc. However, despite DC building on an evolutionary view of resource development, less
attention has been dealt to the evolution of DC over time than one could expect. Partly
this could be due to lacking data (ref), unclear conceptual definitions (ref) or sheer
complexity of studying evolution (ref). One remarkable exception is Zollo \& Winter (2002)
who propose a clear evolutionary argument and hypothesis on DC evolution. Still, little
empirical work has attempted to test their propositions.

\textcolor{red}{DL definition – makes sure all definitions are in order}

By combining insights from 20 years of DC research on external factors such as antecedents
and outcomes with the internal process shaping DC over time, we can shed light on the
workings of DC and extend the knowledge of its implications. Using Zollo \& Winter as our
point of departure we will be able to study the internal process of DC evolution, and by
adding new dimensions related to antecedents and outcomes we will gain insight into
important implications of the concept.

\subsection{Deliberate learning as antecedent and evolution driver}
When Zollo \& Winter (2002) put deliberate learning at the center of their model for DC
evolution they tap into a core tenant of DC theory more developed in later
contributions. Learning theory in organizations distinguishes between autonomous learning
processes \citep{Arthur2005} (e.g. learning by doing \citep{Zollo2002a,Arrow1962} based on
repetitive actions by employees in firms and organizations, and deliberate learning where
learning takes the form of a planned and deliberate process \citep{Arthur2005}. This
deliberation and intent as a driver behind the learning process, corresponds well with the
idea of purposefulness behind strategic intent as evident in the definition of dynamic
capabilities \citep{Helfat2007}: “the ability to {\bf purposefully} build, extend and
modify ordinary capabilities” (our emphasis added). (ref). When organizations invest in
systems to facilitate learning, they invest in capabilities for changing routines in line
with strategic direction of the organization itself. Hence, investments in deliberate
learning corresponds with investments in dynamic capabilities and both concepts contribute
to the development of operational routines.

\emph{H1: Deliberate learning has a direct positive effect on organizations development of
  operating routines.}

They depart conceptually, however, when we look at the functioning of each. The main
function of DL is to accumulate knowledge and experience about the operation of the
organization and its environment (ref). Through this process DL generates insights into
what should be changed and not in order to achieve improved organizational
performance. Contrary, DC deals with how change in operational routines should be done
through putting experience from DL to work (ref). Thus, for DC to be well functioning in
its quest to change OR, a DL process needs to be in place. Without a proper DL the
learning will be less purposeful and hence less attune to the strategic intent of the
firm. Change may still appear, but as organizational change rather than strategic (helfat
distinction between strategix and organizational change). In other words, without DL firms
can achieve technical fitness (ref), but only through DL shaping DC can firms achieve
evolutionary fitness (ref). Consequently, DL is an important antecedent to DC.

\emph{H2: Deliberate learning leads to stronger dynamic capabilities and has a positive
  influence on the level of dynamic capabilities within the firm.}

\emph{H3: Deliberate learning has an indirect effect on operational routines through
  dynamic capabilities.}

Besides impacting the level of DC as an antecedent, DL also contributes to the evolution
of DC over time. Conventional insights suggest that DC cannot be aquired in factor markets
and thus has to be developed internally \citep{Coen2011,Schilke2014a}\todo{(Helfat,
  Finkelstein, Mitchell, Peteraf, Singh,\& Teece, 2007; Katkalo, Pitelis, \& Teece, 2010;
  Maritan \& Peteraf, 2011; Teece, 2014). in Schilke 2014} . This has performance
implications related to the VRIO principle \citep{Barney1991} (as discussed later in this
article), as well as consequences for the evolution of DC. Considering DC as itself a
hierarchy of capabilities \citep{Collis1994,Winter2003} opens up for a distinction between
first-order and second-order DC. On the one hand, systematic learning generates insights
into how routines and lower level capabilities should change (i.e. impacting first-order
DC) \citep{Mahoney1995,Schilke2014a}. On the other, the same learning process is very much
involved in developing and shaping these capabilities over time (Kianto and Ritala 2010;
Schilke2014). Consequently, in addition to being an antecedent to DC, DL also would work
as a second-order DC influencing the evolution of DC over time (the ability of
“learning-to-learn” (Collis, 1994)).

\emph{H4: Deliberate learning has a direct positive effect on changes in dynamic
  capabilities over time}

Dynamic capabilities enables firms to adapt better to changing environments
\citep{Teece1997} and hence creating strategic change in terms of purposefully changing
operational routines \citep{Helfat2007}. Thus, firms with well-functioning DC would be
able to utilize more instances of change and thus more opportunities for learning
(ref). Consequently, high DC leads to more opportunities for both ad-hoc and deliberate
learning to take place and disclose the demand for investments in methods for systematic
experiential learning. As deliberate learning emphasizes a purposefulness of learning by
focusing “on the effect of planned processes for acquiring, codifying, and transferring
existing knowledge within and between organizations” \citep[p. 1160]{Arthur2005}, many
opportunities generated from enhanced DC should lead to more investments in learning
systems. This notion is also supported by previous work emphasizing that doing more of the
same over time leads experience accumulation and hence improved outcomes
\citep{Zollo2004a}. Similarly, by being active in changing its resource base firms
accumulate experience and the importance of deliberate approaches to leading will become
more apparent. A related point is that when the number of opportunities for learning
increases due to higher DC, the incentives for the firm to invest in DL increases due to
potential intensified agency problems \citep{Arthur2005}. The more potential for
knowledge accumulation through learning, the higher potential for misaligned incentives
between workers keeping information back in fear of not being sufficiently compensated,
and management. Hence, the incentive for the firm to invest in DL to better align these
incentives, increases (ibid). In sum these arguments shape an interesting hypothesis on
the feedback loop between DL and DC and how DL could be an outcome as well as an
antecedent of DC \cite{Schilke2018}.(HERE WE ALSO HAVE A DIRECT QUOTE).

\emph{H5: Dynamic capabilities enhances demand for systematic learning so investments in
  deliberate learning abilities will increase over time.}

\subsection{Direct and indirect outcomes of dynamic capabilities}
The objective of strategic change is to enhance a firm’s competitive advantage
(ref). Doing so over time makes for the evolutionary fitness of the firm \citep{Helfat2007}. When firms
are developing capabilities for strategic change through creating, extending and modifying
the resource base this means making decisions to develop operational routines. Such
routines can in turn stem from an inherent learning process evolutionary distinct to the
firm. OR can thus lead to competitive advantage because of their evolutionary character
\citep{Nelson1982,Winter2003}. To the extent OR stems from idiosyncratic
evolutionary process of individual firms rather than being the result of investments in
factor markets, they can satisfy the VRIO condition and thus be a source of competitive
advantage. In such instances, DC works through OR in creating CA. This indirect impact on
CA is very much at the core of the DC literature (ref). Firms with high DC will be able to
better change operating routines to facilitate strategic change in response to changing
conditions facing the firm. DC then creates VRIO OR and hence CA. When this happens, firms
improves their OR and the OR itself satisfies the VRIO condition.

\emph{H6a: Operating routines mediates the effect of dynamic capabilities in generating
competitive advantage.}

On the other hand, resources and capabilities can to a certain extent be captured in
factor markets through investment decisions (ref) or through alliances (ref) or M\&A
activates (ref). By acquiring resources through investments, the operating routines of the
firms may not be directly affected. When these acquisitions happen as a response to
systematic resource orchestration (ref), i.e. reshaping the resource mix to obtain
coordinated resource deployments \citep{Kor2005,Pan2006}, they would indeed have
potential performance implications for the firm. Moreover, when this process of resource
acquisition process of DC develops a way to acquire and mix resource for the purpose of
adapting to change, the resources themselves would not necessarily satisfy the VRIO
condition. The way in which this happens may be much harder to imitate due to the
complexities such routines exhibit. Although resources themselves exhibit clear
equifinality in their impact on competitive advantage, the complex routines of the
orchestration (i.e. dynamic capabilities) would likely satisfy the VRIO condition. In
other words, there are several ways in which resources can be orchestrated to generate CA,
whereas improved OR will always mean improved competitiveness et ceteris paribus. In such
instances DC itself can lead to CA directly.

\emph{H6b: Dynamic capabilities has a direct effect on competitive advantage through the process
of resource orchestration and acquisition of resources in factor markets.}

In other words, we would expect DC to have both a direct and an indirect impact on
CA. Although both the direct and the indirect path stems from DC creating, extending and
modifying the resource base (OR or factor market resources), the difference in the path’s
stems from to what extent the mediating construct can itself satisfy the VRIO
condition. OR can, due to their evolutionary nature, in turn shaped by DC, achieve CA
\citep{Collis1994,Winter2003}. Factor market resources and market acquired capabilities and
best practices can not \citep{Eisenhardt2000,Peteraf2013}.

\begin{figure}

\begin{minipage}[t]{.5\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{./figure/fig1.png}
  \captionsetup{width=0.9\linewidth}
  \captionof{figure}{Conceptual relationship based on Zollo and Winter. Extensions in grey}
  \label{fig:fig1}
\end{minipage}%
\begin{minipage}[t]{.5\textwidth}
  \centering
  \captionsetup{width=0.98\linewidth}
  \includegraphics[width=0.98\linewidth]{./figure/fig2.png}
  \captionof{figure}{Dynamic capabilities direct and indirect paths to competitive advantage determined by the VRIO condition}
  \label{fig:fig2}
\end{minipage}
\end{figure}

Figure \ref{fig:fig1} sums up the hypothesis proposed in this section of our paper. The
white boxes and black arrows represent the parts of the model that was originally proposed
by \cite{Zollo2002a} whereas the grey parts are extensions based on more recent
contributions to the field of dynamic capabilities. Figure \ref{fig:fig2} illustrates the
two proposed paths through which DC affect competitive advantage. The main take away from
this is that dynamic capabilities can affect CA through changing operational routines that
by themselves can satisfy the VRIO condition by means of their evolutionary nature. Thus,
we hypothesize that operating routines mediates the relationship between DC and CA – the
indirect path. Conversely, the same DC enable the firm to identify resources and
capabilities in factor markets that are less idiosyncratic and often takes the shape of
best practices that are definitely imitable and thus bot VRIO. In such instances, the VRIO
condition is satisfied by the DC itself through its complex process of orchestrating and
mixing resources – the direct path.

\section{Data and methods}
An empirical test of the Zollo \& Winter model necessitates a certain time span of
observations in order to enable the study of dynamics over time. The lack of such
longitudinal studies to investigate dynamics has also been a permanent limitation for
moving the DC literature forward empirically and theoretically \cite{Schilke2018}. To
remedy this gap in the literature and to investigate the empirical characteristics of the
Zollo \& Winter model we sampled a subset of Norwegian technology firms ($N=241$) over two
time periods.

We collected survey data from R\&D active firms in Norway at two points in time ($T_1=2005$
and $T_2=2014$). The population was all businesses registered to a scheme for tax deduction
of R\&D costs (called SkatteFUNN). As all enterprises which are eligible for taxation could
register their R\&D activities to receive a tax refund, the registered enterprises include
close to all enterprises which are involved in such activities at the time of our study.

All firms that registered R\&D activities during May to December 2005 were approached
totaling in all 1721 firms. A web-based questionnaire was developed containing the
measures of the key constructs of this paper. A link to the questionnaire was e-mailed to
the firms within a month after they registered R\&D activities. The initial mailing was
followed by two e-mail reminders. Of the firms approached, 1199 (70 \%) returned filled-in
questionnaires. The 1199 companies that filled out the questionnaire were contacted again
9 years later. The majority of firms were contacted in spring/summer 2014. All received a
web-based questionnaire containing the same measures of DC, CA, OR and so on. 283 of the
firms returned filled-in questionnaire.

All the firms in the sample was identifiable by means of the official firm
identifier. This enabled us to attach financial data from the annual accounts of the firms
despite none of the firms in the sample being publicly traded companies. We obtained the
financial accounts from the National Firm Registry (BRREG) for all years between 2007 and
2014 (prior to 2007 was not accessible through our database).

To control for survivor and non-response among our firms over time due to exogenous
factors such as exit, bankruptcy and so on, we estimate a model Heckman correction (Hair,
2010). Applied to the context of this paper, Heckman regression is a method that in the
first step analyses the probability that a firm at $T_1$ will answer the survey at $T_2$
depending on observables at $T_1$. This is modeled as a probit regression. Based on the
probit regression in $T_1$, the inverse Mills ratio (sometimes called lambda) is added to the
analysis in the second step where the goal is to analyze the relationship between our
focal constructs (using regular OLS regression). The inverse Mills ratio is a control
factor which controls for unobserved heterogeneity. More concretely, the inverse Mills
ratio will control for unobserved factors that are related to both firms’ probability of
answering the survey at $T_2$ (ref).

\subsection{Variable construction}

The focal constructs of this paper are deliberate learning, dynamic capabilities,
operational routines and competitive advantage. These are all operationalized using the
survey data described above. Table 1 below describes the items used to operationalize the
focal constructs and their internal validity in forming an additive index (Cronbach´s
alpha). The results show that our key constructs can be represented by the items in the
survey.

There are no established measurement models of dynamic capabilities
\citep{McKelvie2009,Schilke2018}. Therefore this study builds on a mixture of qualitative
case study methodology, literature review and statistical techniques to develop and refine
measures of DCs. First, exploratory qualitative interviews, using a semistructured
interview guide, were conducted with management representatives from 10 R\&D/innovative
firms. The aim of the interviews was to get an overview of each firm's innovation and
development processes, in particular the processes related to its dynamic capabilities and
the management of its resources. Themes raised in the interviews were about network,
co-operation with external R\&D-institutions, learning in the firm, adaptation and changes
in the firm. We interviewed SMEs and larger firms, and the industries varied from
high-tech and ICT to publishing.  Based on the interviews and an extensive literature
review, statements identified as descriptions of dynamic capabilities and resources were
developed and included in a questionnaire.

Second, the informants from the 10 firms were subsequently asked to take part in a pretest
of the questionnaire, including the preliminary items developed to measure dynamic
capabilities and resources, by responding to the questionnaire and giving comments on the
individual questions. This was followed up by a telephone call to the interviewees where
they were asked to report their views on the various questions/items.

Third, the face validity of the items was further examined by pre-testing the measurements
among experts. Researchers with knowledge of business strategy within firms were asked to
evaluate the questionnaire. Based on the results from the pilot study, the items were
adjusted and refined.

Although we recognize at the outset that the concept of dynamic capabilities and their
underlying resource components are very challenging to research in a systematic and
econometric fashion \citep{McKelvie2009}, we follow the argument in the
literature that more empirical work is necessary to test and refine the dynamic
capabilities concept and how it is related to the evolutionary economic theory
\citep{Arend2009,McKelvie2009} \todo{Rosenblom, 2000; Verona and Ravasi, 2003;}.  It
is in this spirit that the research reported in this paper has been undertaken.

Items measuring DC and OCs were developed as statements for which the respondents were
asked to indicate to what extent each statement fitted a description of their business. We
adopted one-sided seven point Likert scale where: 1 = strongly disagree and 7 = strongly
agree. We built on prior studies where items measuring OCs have been measured relative to
competitors \citep{McKelvie2009}.

\text{red}{CONTINUE WITH THE OTHER KEY CONSTRUCTS – INPUT FROM TOMMY}

\subsection{Control variables}

We add a set of control variables expected to explain parts of the variation in our focal
constructs. From the survey we add firm size, firm age, the level of dynamism facing the
firm. Size is the reported number of employees in the firm at $T_2$, and age is the number of
years since establishment. Dynamism is captured by asking a number of questions about the
competitive environment facing the firm. A full list of items and their internal validity
(0.75) is available in appendix ##.

To partly remedy the problem of common method bias, we add controls of financial data from
the firm´s profit and loss statements, and balance sheet. Specifically, we add the natural
logarithm of the assets and debt of the firm to capture the leverage (debt risk) of the
firm, as well as the capital intensity.

One core critique of the dynamic capabilities literature is that it simply is a response
to the firm being a good firm \citep{Arend2009}. This tautology makes it hard to
distinguish the direction of the effects the performance implications of dynamic
capabilities. Simply put both firm performance and dynamic capabilities can be due to some
unobserved factor of the firm simply being a “good firm”. In such cases we would simply
regress one on the other and get positive results without the exact mechanism being
pointed out. To partly control for this we add a proxy for the quality of the firm as a
control variable. In this paper we use the profit margin of the firm as this type of
control. It is not perfect as profit margins can be due to many factors outside our model,
but it serves to control for parts of the possible cofounding factors of firm quality.

\subsection{Estimation and empirical model}

To test our hypotheses, we run a sequential linear regression with heteroskedastic robust
standard errors and Heckmann two-step correction to control for attrition and survival
bias (ref). We test the mediation hypotheses within this framework using a causal
mediation \citep{Baron1986,MacKinnon2008} with quasi-Bayesian confidence intervals
over a million simulations \citep{Imai2010}. The sequential regression
model is presented chronologically in accordance with figure \ref{fig:fig1}. This entails that the
dependent variable in one model will be the independent in the next to study the evolution
of the proposed model sufficiently. The following models are estimated:

\par\noindent
\begin{align}
  OR_{2,i} &= \beta_0 + \beta_1 OR_{1,i} + \beta_2 DL_{i} + \mathbf{\Gamma C_i} + \epsilon_i\\
  OR_{2,i} &= \beta_0 + \beta_1 OR_{1,i} + \beta_2 DL_{i} + \beta_3 DC_{1,i} + \beta_4
               \Delta DC + \mathbf{\Gamma C_i} + \epsilon_i\\
  DC_{2,i} &= \gamma_0 + \gamma_1 DL_{1} + \gamma_2 DC_{1,i} + \mathbf{\Gamma C_i} + \upsilon_i\\
  CA_{2,i} &= \lambda_0 + \lambda_1 CA_{1,i} +
             \lambda_2 DC_{1,i} + \lambda_3 \Delta DC_i +\mathbf{\Gamma C_i}+ \mu_i \\
  CA_{2,i} &= \lambda_0 + \lambda_1 CA_{1,i} + \lambda_2 OR_{1,i} + \lambda_3 \Delta OR_i
             \lambda_4 DC_{1,i} + \lambda_5 \Delta DC_i +\mathbf{\Gamma C_i}+ \mu_i \\
  DL_{i} &= \xi_0 + \xi_1 DC_{1,i} +\mathbf{\Gamma C_i}+ \nu_i
\end{align}

Here $ \mathbf{\Gamma C_i}$ represents a vector of control variables including size, age,
dynamism, financial assets and profit margin. $\Delta DC_i$ is the change in $DC$ from
$T_1$ to $T_2$. Similarly $\Delta OR_i$ represents the change in $OR$.

In addition we estimate mediating models and test formally for mediation using the Baron
\& Kenny method \citep{Baron1986,MacKinnon2008}. We estimate two mediating model to test
if OR mediates the effect of DC on CA, and if DC mediates the effect of DL on OR,
respectively. The mediation equations for OR and DC respectively are state below:

\begin{align}
 \Delta DC_i &= \alpha_0 + \alpha_1 OR_{1,i} + \alpha_2 DL_{i} + \mathbf{\Gamma C_i} + u_i \\
 \Delta OR_i &= \omega_0 + \omega_1 CA_{1,i} + \omega_2 \Delta DC_{i} + \omega_3 CA_{1,i} + \mathbf{\Gamma C_i} + v_i
\end{align}

Finally, we run a Heckman selection model with selecting in to the sample at time $T_2$ is
determined by the following model:

\begin{align}
  P(R=1) &= \eta_0 + \eta_1 L_i + \eta_2 DC_{1,i} + \mathbf{\Gamma C_i} + m_i
\end{align}

$R$ is an indicator variable equal to 1 if the firm has responded at $T_2$ and 0 if
not. $L_i$ is an indicator set to 1 if the person answering the survey is the managing
director or CEO. This acts as an instrument for considering the probability of the firm
selection into the sample, but without having any theorized relationship with the main
outcome variables $DC$, $OR$, and $CA$. It is reasonable to assume that if the CEO was
involved in answering the survey at $T_1$, would increase the probability of the firm
answering at $T_2$ as well.

\section{Results}
The focal variables of our study is presented with pairwise correlation, mean and standard
deviation in table \ref{tab:desc}.It is worth noting that the mean value of OR and DC is
declining over time suggesting that the average firm experience a negative development in
core resources. The ability of the firms in our sample to generate CA is somewhat
unchanged over the sample period. We also observe a certain correlation in the development
in DC over time suggesting that prior period DC is a strong predictor of future
DC. Furthermore, we see a strong correlation between DL and DC which indicates that these
two constructs are indeed related.

\begin{table}[t]
\begin{singlespace}
  \caption{\small Descriptive statistics and correlations for variables in this paper}
  \label{tab:desc}

<<resultscorr, results="asis",echo=FALSE >>=

print(xtable(art2cor ,size="\\small",caption="Correlation matrix \\label{tab:cor}", align = c("l", rep("p{0.3in}",9)),table.placement="h", size = "\\small", tabular.environment="longtable"), size="scriptsize", label='tab:cor', caption.placement="top", floating=FALSE,latex.environments=NULL)

@
\vspace{1ex}

     \raggedright {\scriptsize Notes: Pairwise correlations between variables in the study
    excluding controls}
\end{singlespace}
\end{table}

\begin{table*}[t]
  \centering

\caption{Regression -How are MDC influencing evolution of lower level capabilities }

<<result1, results="asis", echo=FALSE >>=
stargazer(mainres,
          add.lines =list(c("Adjusted $R^2$",r1,r2,r3,r4,r5,r6)),
          single.row=FALSE, column.sep.width = "1pt", omit.stat=c("LL","ser","f","aic","rsq","adj.rsq"),
          no.space=FALSE,
          dep.var.labels=c("OR_{T_2}", "DC_{T_2}","CA_{T_2}","DL"),
        #  column.labels=c("HECKMAN","OLS","HECKMAN","OLS","HECKMAN","OLS"),
          model.names=FALSE,
          object.names=FALSE,

          ### order=c(1,3,2),

           ## covariate.labels= c("$OR_{T_1}$",
           ##                "$\\Delta OR$",
           ##                "$CA_{T_1}$"
           ##                "$DL$",
           ##                "$DC_{T_1}$",
           ##                "$\\Delta DC$",
           ##                "Firm size",
           ##                "Firm age",
           ##                "Environmental Dynamism",
           ##                "Firm profit margin",
           ##                "Firm assets"
           ##          ),

          title= "Regression results for model C", font.size="scriptsize",float=FALSE
          )

@
\label{tab:reg}
\end{table*}

\begin{figure}[t]

\begin{minipage}[t]{.5\textwidth}

  \resizebox{7cm}{!}{\input{./figure/medanalysis1.tex}}
  \captionof{figure}{Mediation analysis 1}
  \label{fig:fig3}
\end{minipage}%
\begin{minipage}[t]{.5\textwidth}
 \resizebox{7cm}{!}{\input{./figure/medanalysis2.tex}}
  \captionof{figure}{Mediation analysis 2}
  \label{fig:fig4}
\end{minipage}
\end{figure}

Table \ref{tab:reg} shows a summary of the regression models proposed in the previous
chapter. Models 1,2 and 3 are designed to test DL as \emph{antecedent} for $OR$ and
$DC$. Consider model 1 we observe that $DL$ is influencing the level of $OR$ within the
firm ($\beta_2 =$ \Sexpr{mod1_dl}) supporting H1. However, the effect disappears when we
add $DC$ in model 2 suggesting a mediating effect. In other words we see that firms with
initial high $DC$($\beta_3=$ \Sexpr{mod2_dc0}) and increasing $DC$ over time ($\beta_4=$
\Sexpr{mod2_ddc}) also have better $OR$. To control for the mediating effect suggested by
the change in coefficients for $DL$ between models 1 and 2, we performed a causal
mediation analysis as reported in figure \ref{fig:fig3}. This analysis confirms the
indications from models 1 and 2 showing that the direct effect from $DL$ on $OR$ is
insignificant and small (ADE = \Sexpr{medmod1_ade}). Meanwhile, the indirect effect
running through changes in $DC$ over time, is significant yielding a positive and
significant indirect effect of \Sexpr{medmod1_acme}. The proportion of the effect that is
mediated amounts to \Sexpr{medmod1_shmed}. In sum these results suggest that H1 is
unsupported, and that the bulk of the effect is indirect running through $DC$, in line with
the expectations in H3. Similarly, we see from model 3 that $DC$ is indeed influenced by $DL$
above and beyond its own evolutionary trajectory. This is in line with our expectations
and lend support to H4.

To evaluate H5 proposing a feedback effect from $DC$ to $DL$, we turn to model 6 in table
\ref{tab:reg}. Controlled for the usual firm level controls we find a positive effect
($\xi_1=$ \Sexpr{mod6_dc0}) from $DC_{T_1}$ on $DL$ measured at $T_1$. This supports H5
suggesting that firms with more $DC$ is more likely to generate opportunities where it
becomes more prevalent that learning will need to take place. One interesting finding is
also that this effect environmental dynamism decreases firms level of $DL$. This is indeed
puzzling and we will look in to this in a post-hoc analysis.\todo{When adding interactions
  we see that high environmental dynamism in conjunction with DC is positive for DL
  meaning that DC will indeed work to make firms more aware of the benefits of DL when
  environments are changing} It is also worth noting that models 1,2 and 3 have positive
and significant $\rho$ from the Heckman selection model indicating that selection is
leading to biased estimates. By running a Heckman correction we avoid this potential
problem.

Models 4 and 5 in table \ref{tab:reg} deals with the \emph{outcome} of DC. H6a and H6b
both considers effects of $DC$ on the competitive advantage of the firm $CA$. $CA$ is
influenced by the level of $DC$ of the firm indicating that dynamic capabilities are a
source of competitive advantage in line with our expectations. Furthermore, considering
model 5 we see that $OR$ also has a significant positive effect on $CA$ indicating that
operating routines can also be a source of advantage for the firm. Moreover, the
coefficient of $\Delta DC$ decline from \Sexpr{mod4_ddc} to \Sexpr{mod5_ddc} when we add
$OR$. This suggests that some mediation is at work, i.e. that changes in $DC$ leads to
changes in changes in $OR$ leading to $CA$. This indirect path is in line with the
expectations from our hypotheses and as described in figure \ref{fig:fig2}. When formally
testing this mediation running a causal mediation analysis (see figure \ref{fig:fig4}), we
find added support that both a \emph{direct-} and an \emph{indirect} path exist to explain
the relationship between $DC$ and $CA$ (ADE = \Sexpr{medmod5_ade}, ACME =
\Sexpr{medmod5_acme}) and a total mediated share of \Sexpr{medmod5_shmed}.

Table \ref{tab:hyptab} depict a summary of the hypotheses and indicate if they are
supported or not in our analysis. We will now turn to the discussion.

\begin{table}
     \footnotesize
  \centering
\input{./figure/hyptab.tex}}
\end{table}

\section{Discussion}

The objective of this article is to study the antecedents and outcomes of dynamic
capabilities, particularly the role of deliberate learning in shaping dynamic capabilities
over time. Learning is arguably an important determinant of firm success (ref arrow and
more), and the deliberate investment in learning to learning (i.e. deliberate learning)
has a long standing position as a theoretical antecedent of dynamic capabilities
\citep{Zollo2002a,Eriksson2014,Schilke2018}. Furthermore, the role of deliberate learning
in creating better operational routines is well established (ref), but it`s role in
generating competietive advantage for the firm, is less understood. Alas, the main
mechanism, the VRIO principle \citep{Barney1991a} is well known and still deemed highly
relevant relevant in understanding the source of competitive advantage of the firm
\citep{Priem2001}. Despite the prevalent and considerable citation of the
\cite{Zollo2002a} model, neither an empirical analysis of the model propositions, nor it`s
connection to outcomes such as competitive advantage has, to our knowledge, been
undertaken.

Our proposed extension of \cite{Zollo2002a} as depicted in figure \ref{fig:fig1}, taps
into the debate on the most important \emph{antecedents} and \emph{outcomes} of dynamic
capabilities \cite{Eriksson2014,Schilke2018}. Our results indicate that deliberate
learning is indeed an important antecedent of dynamic capabilities. Firms with stronger
capabilities for deliberate learning will develop higher levels of dynamic
capabilities. Thus the deliberate decision to invest in systems for learning will lead to
the evolution of better capacities to "purposefully create, extend, or modify its resource
base" \citep[p. ###]{Helfat2007}. Indeed, this also follows from Zollo \& Winter`s
definition where "a dynamic capability is a learned and stable pattern of collective
activity through which the organization systematically generates and modifies its
operating routines in pursuit of improved effectiveness" \citep[p
###]{Zollo2002a}.

The next step of figure \ref{fig:fig1} is to establish the link between deliberate
learning and the operational routines of the firm. These operating routines are what
\cite{Winter2003} calls "how we can earn a living now" (p. 992) or the technical fitness
of the firm \citep{Helfat2007}. Such operational routines are the product of learning
processes over time and should indeed be linked to deliberate learning of the firm as
proposed by \cite{Zollo2002a}. However, we find that this direct effect between deliberate
learning and operational capabilities disappear when we introduce dynamic capabilities into
the model. This suggest that dynamic capabilities are evolving as a response to deliberate
learning and, in turn, shaping operational capabilities in line with expectations of Zollo
\& Winter. In other words, we find support for an indirect link between deliberate
learning and operational routines, but no direct path.

This finding may come from the fact that learning works differently at different levels of
the capabilities hierarchy \citep{Collis1994}. Considering operational capabilities as
zero-order capabilities \citep{Winter2003,Schilke2014a} defined as more "geared towards
the operational functioning of the firm" \citep[p. 340]{Zollo2002a}, learning may be more
ad-hoc and less routinized as experiences are accumulated "on the fly" or through
"learning by doing" (ref) rather than deliberate and systematic. Indeed, learning by doing
is established as a robust result in the empirical literature on it`s link to operational
excellence (ref). When learning becomes deliberate and systematic through investments in
learning to learn capabilities, this would not automatically strengthen the short term
evolution of operational capabilities directly \citep{Arthur2005}. \todo{Need some more
  references here}. These findings would also lend support to the notion of higher-order
dynamic capabilities (capabilities to change dynamic capabilities)
\citep{Collis1994,Winter2003,Schilke2014a} and that deliberate learning could be
classified as such. In contrast, ad-hoc learning based on repetitive actions by employees
does not satisfy the requirements for a dynamic capability as the actions are less
strategic and purposeful.

An interesting finding is also that deliberate learning itself could be seen as a function
of dynamic capabilities. We argue that this stems from firms with high levels and
evolution of dynamic capabilities will take more advantage of adapting to change and hence
creating more opportunities for learning. This leads the firm to realize the need for
more deliberate approach to systematic learning. Particularly, this would lead to more
investments in systems for learning and, in turn, improvements in dynamic and operational
capabilities. For obvious reasons this is a tough causal claim to make and we have little
support for a causal argument. The insight will thus need considerably better data and
empirical analysis going forward. However, it has some support in the literature where
learning can be seen as an outcome as well as an antecedent to dynamic capabilities
\citep{Schilke2018}. \todo{maybe add someting about the interaction with dynamism}

Moving from antecedents to outcomes we extended the \cite{Zollo2002a} model by adding
competitive advantage to the model. We find that both operational capabilities and dynamic
capabilities can lead to competitive advantage. The most obvious path from the mere
definition of dynamic capabilities, runs through the evolution of operational
capabilities. Dynamic capabilities shape operational capabilities and changes them to
adapt, hence safeguarding that they are adapted. This is beyond the ad-hoc learning and
adaptation taking place as repetitive actions by employees. TO BE CONTINUED

\subsection{Limitations}

Our operational construst of $DC$ captures dynamic capabilities in
a broad, non-functional, sense (e.g. as a broad-purpose capability rather than one for
a particular purpose) \citep{Schilke2018}, and will hence capture a mind-set of the
organization rather than a narrow set of learned capabilites. This makes our conclusions
more general, albeit maybe less precise in predicting what investments in deliberate
learning would nmean for important capacities such as new product development and
innovation. 
 BLABLABLABLA

\section{Bibliography}

\singlespacing
\bibliography{Dissertation_clean}

\section{Appendix}

\end{document}