### 2019-10-14 article4_tommylars.rnw
still testint

%https://www.gnu.org/software/emacs/manual/html_node/emacs/Fill-Commands.html
%https://academic.oup.com/journals/pages/authors/latex_files
%latexmk -pdf -pvc -auxdir=/Users/larshovdanmolden/Documents/Output/  article2_mancap.tex

\documentclass[review,fleqn]{elsarticle}
\usepackage[textsize=tiny,colorinlistoftodos]{todonotes}
\usepackage{subfig}
\usepackage{subcaption}

\input{preamble.tex}
%%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\setlength{\mathindent}{0pt} %% NB MAYBEREMOVE BEFORE SUBMISSION

%%%% TURN OFF PAGE NUMBER FOR INCLUSION IN THESIS
%\pagenumbering{gobble}
%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Dynamic capabilities and competitive advantage}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
%\author{Lars Hovdan Molden \fnref{myfootnote}}
%\address{Kongensgt 42, 7713 Steinkjer}
%\fntext[myfootnote]{Nord University Business School}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
 We find that DC as routines to shape routines is most prevelant under moderate dynamism,
whereas DC as routines to capture resources in factor markets are more prevelant under
dynamic environments.

\end{abstract}

\begin{keyword}
\texttt{Dynamic Capabilities, Deliberate Learning, Dynamics }
\end{keyword}

\end{frontmatter}

%\linenumbers

\doublespacing

\section{Introduction}

%% Extant literature on dynamic capabilities is battling with a fundamental conception and
%% prediction of the theory, namely if, when and how dynamic capabilities can lead to competitive
%% advantage.

Seminal contributions to the field of dynamic capabilities have argued for both direct
\cite{Kor2005,Eriksson2014,Schilke2014a} and indirect (ref) effects of dynamic
capabilities on competitive advantage, although increasingly the bulk of the literature
focuses on the latter \citep{Schilke2018}. This indirect link is typically understood as
dynamic capabilities modifying, extending or creating firm resources (helfat) (or
lower-level operational capabilities (Winter, collis)) and thus generating competetive
advantage (Eriksson and others). Furthermore, empirical evidence keep suggesting that both direct (ref)
and indirect (ref) links exist. This warrant the question of \emph{when} direct and
indirect links between dynamic capabilities and competitive advantage is most pronounced.

Moreover, firm resources or capabiltiies themselves are far from a homogenous set and
varies among other things with respect to their uniqueness, inimatbility and value
(barney). Important contributions have thus argued for a deeper insighi into these
intermediaries and their mechanism calling for a "deeper insight into the variety of mechanisms that
underlie the performance effects of capabilities.” (Schilke, 2014a: 199). 
as well as \emph{what} different intermediaries exist.

This paper investigates the boundary conditions of the VRIO condition and how dynamic
capabilities and resources interact to create firm competitive advantage. We ask the
following research question:
Under which conditions are dynamic capabilities able to yield competietive advantage
through resource orchestration and what are the different outcomes of different types of resources.

We find thatthe dynamism surrounding hte firm represent an important boundary condition as
suggested in parts of the literature. However, DC consinutes to generate competetive
advnatage also in high-velocity environments, but doing increasingly so indirectly through
resource orchestration. Furthermore, we find that non-vrio resources are more likely to
yield competetive power in high velocity environments, and that vrio-resources tend to
maintain their CA under any condition.

We demonstrate that two paths indeed do exist and that their relative importance varies
with the degree of environmental dynamism facing the firm.

Under what cionditions are the direct path and indirect path most prevalant? Under which condi

The direct

\section{Hypotheses}\label{sec:hyp}

%% We can use the drive train to determine hwo DC works differently under different levels of
%% dynamism. It is not that DC does not work under stable conditions as EM claims. Is only
%% that the effect runs stringer through the evolutionary path. Whereas the market path is
%% stringer under higher dynamism

The link between firm resources and competitive advantage is determined by the extent to
which resources are valuable, rare, inimitable and organized (VRIO)
\citep{Barney1991a}. This principle is also used to evaluate how higher order capabilities
(e.g. dynamic capabilities \citep{Winter2003,Collis1994}. This is the core principle we
use in this article to evaluate how dynamic capabilities can and cannot create competitive
advantage for the focal firm. Furthermore, as both higher order capabilities (e.g. dynamic
capabilities) as well as lower-order resources and capabilities (operating routines and
capabilities) can and cannot satisfy the VRIO condition, we find it useful to think of a
multilevel VRIO structure in firms.

\subsection{Dynamic capabilities and VRIO}

One contentious issue surrounding the DC theory is the extent to which differing views on
the nature of dynamic capabilities \citep{Peteraf2013} as well as their outcome. One
strand of contributions sees DC as routines that are simple in nature and more as best
practices to consider (EM). Thus, dynamic capabilities are more like behavioral patterns
and less idiosyncratic to the firm and will consequently not be able to satisfy the VRIO
condition for competitive advantage to emerge. Dynamic capabilities as such routines are
short lived, unstable and “substitutable” (EM: 1110) thus violating a key VRIN condition,
and they are more “more homogeneous (..) than is usually assumed” (EM: 1116). However, the
simplicity of simple rules also makes them faster do deploy and put in motion, albeit
potentially unstable. To the extent dynamic capabilities manifest themselves as simple
rules they cannot alone create competitive advantage, but could very well generate
strategic change by facilitating the process of resource orchestration. Resource
orchestration is defined as "blablbalbla" (helfat). In other words, even if DC play out as
simple rules that are not in themselves satisfying the VRIO condition, the way they work
through creating, extending and modifying firm resources, can still yield VRIO results.

On the other hand, \cite{Teece1997} and others \cite{Helfat2007} argue that dynamic
capabiltiies will rather invariably be VRIO due to the evolutionary nature of their
development \citep{Arndt2018,Peteraf2013}. Arguing that dynamic capabilities reflect more
complex processes that co-evolve with responses to market changes over time and hence
creating idiosyncratic capabilities that are unique, valuable and hard to copy
\citep{Helfat2007,Arndt2018}. Consequently, they are likely to satisfy the VRIO condition
and thus leading to competitive advantage.

Increasingly the literature remain focused on an \emph{indirect
  path} between dynamic capabilities and competitive advantage
\citep{Barreto2010,Schilke2018}. The theory behind this path goes back to the core of
dynamic capabilities where resource orchestration happens as a response to changes in the
environment. When markets, technologies or other external factors changes, firms with high
levels of dynamic capabilities are able to seize opportunities for changes, seize these
opportunities and transform the resource base to fit the strategic change needed to stay
competitive \cite{Teece2007}. This ability to adapt is by some coined \emph{evolutionary
  fitness} defined as BLABLBALBA and stems from the ability of the firm to reconfigure
resources in the face of change. The extant literature, however, is less clear on how the
nature of the underlying resources matter in this indirect path. On the one hand we can
see

Following REF suggesting more reseaarch on dynamic capabilities should put forth
competeing hypotheses we propose these first

We introduce the notion of multilevel VRIO to integrate the interactions between different
types of capabilities at different levels. This entails that dynamic capabilities (the
high level) can consist of complex and evolutionary characteristics that are idiosyncratic
to the firm and hence VRIO by definition along the lines of how the concept is understood
by \cite{Teece1997}. On the other hand dynamic capabilities can also take the shape of
simpler routines which are short lived, unstable and substitutable thus violating a key
VRIO condition. This follows the notion of dynamic capabilities found in
h\cite{Eisenhardt2000} arguing that they are more “more homogeneous (..) than is usually
assumed” (EM: 1116).

Moreover, these dynamic capabilities are in the extant literature argued and found to work
directly \citep{Kor2005,Eriksson2014,Schilke2018} on competitive advantage which suggests
that they are indeed VRIO. On the other hand, a growing body of literature (ref) as well
a consensus on the definition of dynamic capabilities (ref), holds that they can lead to
competitive advantage by changing underlying resources and routines. In this
\emph{indirect path} the dynamic capabilities itself need not be VRIO as long as the
underlying resources themselves are.

\subsection{The role of dynamism}

The most contentious issue in the current debate on dynamic capabilities is how their
nature varies with the environment.

If htey are indeed latent

The lower level resources and routines are to be understood as conventional resources in
line with the resources-based view

WHAT IS THE ROLE OF DIRECT VS INDIRECT, HOW DOES IT RELATE TO THE ENVIRONMENT, AND WHAT
ROLE DOES THE TYPE OF UNDERLYING RESOURCE PLAY

Conventional dc literature has argued for both indirect and direct links. Indeed both
have been found, but while the indirect link is well rooted in the theory, the direct link
is to our knowledge less clear. The core idea of DC is that it works to modify underlying
resources as a response to changes in the environment and hence creating strategic change
needed to stay competitive. This resource orchestration is heavily represented in the
literature (ref) and through this process firms are able to generate CA. In such instances
the CA can stem from VRIO resources and/or VRIO DC. A VRIO DC would mean that the firm is able
to shuffle and reorganize generic resources in a way that creates a new and unique
configuration of resoures idiosyncratic and valuable to the firm. In such instances, the
outcome whould be CA. On the other hand, the DC may not by itself be VRIO but rather a set
of simple routines put in motion to change, but if the resources they reconfigure are
themselves VRIO, the outcome could just as eqsily be a competetive advantage. this
equfinaity is discussed in the literaure \cite{DiStefano2010} and makes it hard to
distingusih between the role of DC and ordinary capabilites in generating CA.

Furtemore, Considering a direct link between DC and CA, DC it self must consist of some
VRIO characteristics to garner CA. So in owther words, when is DC working directly and
indirectly_

How are dynamic capabilities working directly on competitive advantage? The microfoundations of dynamic capabilities explicated by \cite{Teece2007} is a useful
point of departure of this theoretical inquiry. He hold that for "analytical purposes,
dynamic capabilities can be disaggregated into the capacity (1) to sense and shape
opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness
through enhancing, combining, protecting, and, when necessary, reconfiguring the business
enterprise’s intangible and tangible assets". Sensing and seizing is a part of what we
would coin the "opportunity capture" of dynamic capabiliites, where as reconfiguring
concerns itself with the "resource orchestration" needed to change assets to stay
competetive.

Hence, when considering indirect and direct paths to compettetive advantage we will have
to be clear on two fronts: Are DC themselves VRIO and are the underlying resources and
routines central to the resource orchestration, VRIO.

       Path   LOW     t  HIGH     t
1 DC -> CA1 0.290 2.732 0.093 0.996
2   DC -> P 0.325 2.896 0.339 3.284
3  P -> CA1 0.064 0.462 0.171 1.913
>

The microfoundations of dynamic capabilities—the distinct skills, processes, procedures, orga- nizational structures, decision rules, and disciplines—which undergird enterprise-level sensing, seizing, and reconfiguring capacities are difficult to develop and deploy.
For analytical purposes, dynamic capabilities can be disaggregated into the capacity (1) to sense and shape opportunities and threats, (2) to seize opportunities, and (3) to maintain competitiveness through enhancing, combining, protecting, and, when necessary, reconfiguring the business enter- prise’s intangible and tangible assets.

t its most fundamental level innovation research concerns itself with a simple
input-output model \citep{Adams2006} where innovation output is seen as a function of
efforts put into the innovation process. However, firms vary with respect to their ability
to utilize input and transform it into innovation output. This transformation of input
into output has been studied with different perspectives and conceptualizations such as
productivity of R\&D \citep{Gwynne2015}, innovation capability \citep{Breznik2014},
innovation factor productivity (ref) and innovation conversions rate
\citep{Duran2015}. \newline

\subsection{The idiosyncratic nature and strategic expression of IC}
The question of why firms vary in the degrees of innovation output they get from similar
levels of innovation input, leads on to a higher order capablility that coordinates and
extends the first order routines and resources as input in the innovation process. By
innovation input we mean the resources and routines put into the effort creating
innovation which composition is dealt with in a large literature (see \cite{Becheikh2006}
for a review). Innovation output means new products, processes, business models, and
organizational traits as the results of a purposeful enactment of resources (Van de Ven
1995). The mechanisms through which input is transformed to output is a reflection of firm
idiosyncratic abilities. At the meta level this concept is related to entrepreneurial
orientation \citep{Lumpkin1996}, entrepreneurial action \citep{Klein2016}, innovation
capability \citep{Breznik2014}, and absorptive capacity \citep{Cohen1990a}. We go with the
term \emph{innovation capability (IC)} defined as the "ability to continuously transform
knowledge and ideas into new products, processes and systems for the benefit of the firm
and its stakeholders"\cite[p. 384]{Lawson2001}. IC is characterized by some factors
similar for all the above mentioned concepts such as the ability to \emph{sense, seize} and
\emph{transform} opportunities into outcomes \citep{Teece2007}, and it`s idiosyncratic
nature is argued by several earlier contributions \citep{Duran2015,Klein2016}. Or as put by
\cite{Duran2015}: "While most researchers assume that a positive correlation exists
between innovation input and innovation output, the “slope” of this relationship likely
differs among organizations because of the complexity of innovation management." (ibid p
##). Innovation management (def) is the managerial levers for creating innovation
capabilities, i.e. their ability to orchestrate resources \citep{Sirmon2011} which is an
important part of what \cite{Helfat2007} calls \emph{Dynamic Managerial Capabilities
  (MDC)} defined as "the capacity of managers to purposefully create, extend, or modify
the resource base of an organization" (ibid p 24). MDC, and hence IC, is heterogeneous
between firms due to the evolutionary nature of developing such capabilities
\citep{Teece2014,Eriksson2014}. Consequently, we propose the following hypothesis:\newline

H1: IC is heterogeneous between firms due to differences in their sensing and seizing
abilities \newline

\cite{Teece2007,Teece2014} makes a sharp distinction between ordinary- and dynamic
capabilities arguing, among other dimensions, that managerial emphasis in the former
relates to cost control, whereas DC relates to entrepreneurial asset orchestration,
i.e. "doing the things right" in novel ways. A similar line of reasoning argues that
dynamic capabilities differs from ordinary routines and capabilities in that it goes
beyond best practices in operational problem solving \citep{Eisenhardt2000}. Consequently,
we would expect IC to be positively related to innovation outcomes as firms with high IC
would generate more innovation. Moreover, as managers face changing environments, the ones
with strong IC would emphasize innovation activities, whereas those with weaker IC would
have to resort to other means of adapting to change, e.g. cost efficiency. In other words,
under environmental dynamism high-IC companies will engage in sensing, seizing and
transforming opportunities, trusting the evolutionary fitness of
their dynamic capabilities, and preferring innovation as strategic choice. Meanwhile, low-IC
companies will revert to operating, administering and governing the business, trusting
their technical fitness and focus strategic choice on efficiency
\citep{Teece2014}. \newline

H2: IC is positively related to innovation activities \newline
H3: IC is negatively related to efficiency enhancing activities \newline

\subsection{The relative importance of sensing and seizing}

The concept of sensing and seizing has analogous meaning in theoretical traditions tangent
to DC. Notably entrepreneurship \citep{Hitt2011} and organizational theory
\citep{March1991}. \cite{Hitt2011} states that "entrepreneurship involves identifying and
exploiting opportunities" (ibid p 59) where identification is tied to the process of
discovery \citep{Shane2000}. Based on the definitions of opportunity discovery as DEF HERE
juxtaposed to the definition of sensing as "a scanning, creation, learning, and
interpretive activity" \citet[p. 1322]{Teece2007} we argue that they are similar in nature
and usable for exploring the importance of the latter in the innovation process. It also
bears resemblance to the concept of organizational exploration as searching, risk-taking,
discovery, experimentation, prototyping, and flexibility \citep{March1991}. Independent of
the nature of opportunities to be sensed as being either objective
\citep{Alvarez2007,Shane2000} or created \citep{Alvarez2007}, the process of sensing does
require resources and capabilities for search (ref) and a clear affinity between
opportunity and a project exist (casson and Wadeson in Mcmullen2007) suggesting s clear
inter dependency between routines and resources, and higher order capabilities of sensing
\citep{Teece2007} and making sense of the opportunity
\citep{Weick1995,March1991}. Consequently, we would expect the sensing class of dynamic
capabilities to be positively related to innovation on the one hand, and reinforcing the
impact from innovation input simultaneously.\newline

H4: Firms with strong sensing abilities will yield higher innovation output \newline

H5: Sensing as dynamic capability reinforces input in the innovation process to
generate innovation output \newline

In similar ways the concepts of seizing \citep{Teece2007} and exploitation \citep{Shane2000}
bears resemblance. However, by virtue of its nature innovation represent novel processes
within the firm so that parts of the understanding of exploitation is
redundant. Exploitation is defined as utilizing old certainties within the organization
\citep{March1991} - that is existing resources and capabilities. Innovation does by its
very nature represent a shift from existing resources and certainties (albeit often a very
incremental one, but a shift nontheless). Hence the dichotmous nature of exploration vs
exploitation does not always ring true for innovation processes. However, the discovery and
exploration of new opportunities is in any case a first step of any innovation process,
and sensing is the core class of dynamic capabiltieis in place to facilitate this
process. When the opportunity has been discovered through sensing, the firm`s ability to
capture value and generate innovation will rest on its ability to seize the opportunity -
or in other words - exploit this newly arisen opportunity. As such the exploitation
happens not on existing certainties, but of the potential future certainty arising from
the innovation opportunitiy from the sensing. Consequently, we argue that the concept of
seizing as rooted in a sense of grasping opportunities already sensed, is a more useful
construct understanding firm innovation. \newline

Similar to sensing, the process of seizing requires considerable efforts in enhancing
competencies and routines (i.e. orchestrating \citep{Hitt2011}) to grasp the opportunity
identified \citep{Shane2000}.\newline

H6: Firms with strong seizing abilities will yield higher innovation output \newline

H7: Seizing as dynamic capability reinforces input in the innovation process to
generate innovation output \newline

Despite being conceptually sequential (i.e. seizing follows sensing) their relative
importance to the innovation process is less clear from the existing literature. However,
seizing is recognizing potential from the sensed opportunities by combining novel insights
and market evolution and dynamics. Leveraging this opportunity does, however, have
implications for strategic choices and decisions to invest \citep{Teece2007,Helfat2015}
and such pursuits will likely require asset specific investments
\citep{Williamson1985}(MORE TO THE POINT REF HERE). This mechanism is fundamental as it
ties into the incentives to develop such capabilities \emph{ex ante}
\citep{Stieglitz2009}. When faced with a new venture opportunity firms will have to make
timely investment decisions with potential considerable lock in effects leading to a
Stackelberg-type investment problem \citep{Arend2009}. Comparatively, the processes of
sensing can incur with considerable search and discovery costs (REF), but "sensing does
not necessarily involve large commitments of resources, at least not relative to seizing"
\citep[p. 1343]{Teece2007}.Thus, due to its asset specific requirements, the expected
"return" from seizing should be higher than that of sensing.\newline

\noindent H8: Seizing is of relatively stronger importance in transforming innovation
input to output \newline

Finally, turning to innovation output as a first step in generating value, seizing also
plays into the process of generating business models for capturing value from innovation
\citep{Teece2007,Helfat2015}. A business model can be understood as a "pathway by which
technological innovation and knowhow combined with the utilization of tangible and
intangible assets are converted into a stream of profit"
\citep[p. 40]{Teece2017}. Consequently, seizing will play a significant role in
transforming innovation into value for the company by re-orchestrating resources and
capabilities \citep{Teece2010,Teece2017}. Hence, we will see seizing working in different
parts of the input-output process. \newline

\noindent H9: Seizing reinforces the process of generating value from innovation \newline

%\subsection{Balancing sensing and seizing - ambidexterity?}
%\begin{answer}
%  Not really sure if this is important or if it just adds confusion. Perhaps suggesting
%  writing a separate conceptual paper on the differences between exploration /exploitation
%  vs sensing/seising at a later stage
%\end{answer}

%The classical notion of the innovator`s dilemma (ref christensen) holds that
%firms face a trade-off between pursuing new avenues for business
%opportunities on the one hand (novelty), and maintaining efficient organizational
%characteristics on the other (efficiency). This trade-off is arguably leading certain
%firms to inertia in the face of environmental change due to limited abilities to pursue
%both. If this holds true we would expect a firm with considerable abilities for sensing, to be less
%able to seize, and vice versa. However, a related stream of literature argues that firms
%can do both by being \emph{ambidexterous} \citet{oreilly2004}.   \newline

\section{Studying innovation capabilities in context}
There are two important obstacles to empirically studying the working of innovation
capabilities as dynamic capabilities at the firm level. First, dynamic capabilities are
mainly considered valuable and operational in dynamic environments
\citep{Teece1997a,Helfat2007}, thus finding the proper context capturing this dynamism is
crucial. When studying the phenomena across a multitude of industries and firms with
simultaneously, the challenge arise to identify a proper context for inquiry. We
understand environmental dynamism to exhibit volatile and unpredictable characteristics
such as unstable demand conditions and shifts in industry structure, and increased
probability of supply or demand shocks \citep{Sirmon2007,Schilke2014}. In such
environments states, situations, unfamiliar to the firm occur and challenges it`s ability
to make sense of the market \citep{Weick1995}, i.e. matching the environment with a proper
internal response. Moreover, firms meeting new challenges and finding existing routines
and resource configurations to be insufficient in coping, tend to discourage truly novel
explorative solutions \citep{schreyogg2007}. This combined "matching" and "inertia" problem
will be remedied by dynamic capabilities making them a mediating factor for environmental
dynamism \citep{Schilke2014}. In addressing this challenge of context we find the 2008
global financial crisis as a useful back drop. The rapid decline in financial asset value
coupled with soaring volatility, and subsequent falling demand in the real economy,
constitute an interesting backdrop for our study \citep{Makkonen2014}.\newline

A second challenge is the measurement and operationalization of innovation capabilities as
dynamic capabilities \citep{Schilke2014}. A majority of studies are, however, relying on
proxies with single items capturing the nature of dynamic capabilities
\citep{McKelvie2009}, but several studies are increasingly building latent measures based
on observable characteristics to develop a more nuanced understanding of the phenomenon
\citep{Makkonen2014,Schilke2014}. We reconfigure the framework put forth by
\cite{Helfat2015} on how observable traits at the management level is operationalized into
sensing, seizing and transforming classes of dynamic capabilities. Following
\cite{Breznik2014} arguing for the similar nature of dynamic- and innovation capabilities,
we focus on the sensing and seizing aspect as most relevant for studying innovation in an
input-output framework. Sensing and seizing are, as argued earlier in this paper, related
to other well established constructs in the innovation and entrepreneurship literature and
hence suitable for the innovation context, most notably the idea of resource orchestration
\citep{Hitt2011} and exploration/exploitation \citep{March1991}.\newline

\cite{Helfat2015} demonstrate that certain cognitive capabilities at the management level
are antecedents of dynamic capabilities. Managerial \emph{perception} and \emph{attention}
enhance the ability for sensing. Perception relates to sensing in that management "use
cognitive frameworks they possess to “connect the dots” between changes in technology,
demographics, markets, government policies, and other factors"
\citep[p. 831]{Baron2006}. Attention, in turn, are related to perception, and hence to
sensing, as it enables managers to focus on particular information to focus "on relevant
stimuli (..)[to] facilitate environmental scanning" \citep[p. 839]{Helfat2015} necessary for
opportunity discovery \citep{Companys2007} and, consequently, sensing.\newline

Underpinning seizing are capabilities for \emph{problem solving} and
\emph{reasoning}. These are treated as one construct by \cite{Helfat2015}. In our
input-output model it is useful to distinguish between seizing innovation opportunities in
the form of deciding to invest and pursue discovered potential in the sensing process
(hence in helping generating innovation output), and the process of capturing value (i.e
generating innovation gain) through developing business models. \cite{Helfat2015} does not
explicitly make such a distinction presumably because their dependent variable is a
broader "firm performance" and because the impact of problem solving and reasoning is
working similarly in creating innovation output and innovation gain. \newline

Having found a useful context and a conceptual framework for operationalizing innovation
capabilities, we now turn to the data and methods applied in this paper.

       name Mean_est     t
3   DC -> P    0.396 6.157
4   DC -> R    0.416 8.104
5 DC -> CA1    0.239 4.018
8  P -> CA1    0.172 2.724
9  R -> CA1    0.081 1.262
> R
       Path   LOW     t  HIGH     t
1 DC -> CA1 0.417 6.585 0.144 1.513
2   DC -> R 0.388 6.078 0.534 7.525
3  R -> CA1 0.017 0.219 0.211 2.123
> P
       Path   LOW     t  HIGH     t
1 DC -> CA1 0.355 4.805 0.133 1.379
2   DC -> P 0.386 5.294 0.436 4.663
3  P -> CA1 0.155 2.193 0.232 2.315
> B
       name R_LOW_est   t.x R_HIGH_est   t.y
1 DC -> CA1     0.355 5.407      0.044 0.400
2   DC -> P     0.376 3.729      0.432 4.668
3   DC -> R     0.366 5.569      0.522 7.225
4  P -> CA1     0.158 1.989      0.220 2.357
5  R -> CA1     0.010 0.135      0.200 2.033
>

%% creating and One example
%% is the role of post-acquisition routines where routines to identify and deal with
%% acquisition partners are a source of generating strategic change (Eisenhardt and Sull
%% 2001, ref). Another example is (see refs from Schilke 2018). The functioning of this
%% expression of dynamic capabilities are more latent and hence mainly observable ex
%% post. Hence we can see this as direct route to competitive advantage as this expression of
%% dynamic capabilities work directly in the factor market in capturing resources
%% needed for strategic change. We coin this direct connection as the \emph{market path}.

On the other hand, a stream of contributions see dynamic capabilities as evolutionary and
more complex routines (ref) or processes (ref) set to change underlying capabilities in
the face of changing conditions (TPS). This notion stems from evolutionary economics
\citep{Winter2003,Nelson1982} and the behavioral theory of the firm \citep{Cyert1963}
holding that path dependency and complex processes for changing underlying operating
routines are a source of competitive advantage for the firm. Along this indirect path
change is gradual and running through changes in underlying routines. We coin this the
\emph{evolutionary path}.

These paths from dynamic capabilities to competitive advantage are different expressions
of dynamic capabilities with different mediums of intermediation. The \emph{market path}
works in the market and his hence not typically altering lower order capabilities and
resources in the firm directly, but rather acquiring new ones. The source of competitive
advantage along this path stems from the interaction with the other, \emph{evolutionary
  path} in a dynamic bundle \citep{Peteraf2013}. Such interactions create a bundle of
capabilities and routines that are socially complex and hard to imitate
\citep[p. 320]{DiStefano2014}. Moreover, the interactions between the two paths of dynamic
capabilities is separable with respect to what intermediation exist. Consequently,
separate paths exist albeit a part of a dynamic bundle.

\cite{DiStefano2014} propose an integrated framework where simple routines and more
complex processes could contribute simultanously to the firms outcome, specifically its
competitive advantage. In this framework, coined the "organizational drivetrain" the
authors envision the movement of an organization as set in motion by a drivetrain system
not unlike that of a bicycle.

Building on the notion of an organizational drivetrain we argue that dynamic capabilities
can indeed impact the organizational movement or change through two distinct but
interconnected mechanisms (i.e. the crank set and the rear shifter). The two mechanisms
consitutting the drivetrain is consistent with our notion of \emph{evolutionary path} and the \emph{market path}. The
evolutionary path is instigated by more complex processes setting in motiong a chance in
the lower order (operational) routines \cite{Collis1994,Winter2003}. In the drivetrain
metaphor this path corresponds to change stemming from the front crank set of thich there
are fewer gears, but with larger impact. This path is slower
and yields a more idiosyncratic set of organizational routines which is very much in line
with the idea put forth in the seminal founding papers \citep{Teece1997,Winter2003}.

Seeing the bike rider and her efforts to handle the levers of control (i.e. the gear
shifter) as the dynamic capability, and consider it not only as a management effort, but
that of the organization as a whole. The process of changing front gears are more profound
(i.e. evolutionary) and costly, and it is for larger and slower strategic
changes. Operating with simple routines in factor markets, however, is quicker and the
rider does this on a more frequent basis with less effort.

H1: Dynamic capabilities works to create competitive advantage through two distinctly
different paths.

\subsection{The two paths of dynamic capabilities}
The objective of strategic change is to enhance a firm’s competitive advantage
(ref). Doing so over time makes for the evolutionary fitness of the firm
\citep{Helfat2007}. When firms are developing capabilities for strategic change through
creating, extending and modifying the resource base this means making decisions to develop
operational routines. Such routines can in turn stem from an inherent learning process
evolutionary distinct to the firm. Operating routines can thus lead to competitive
advantage because of their evolutionary character \citep{Nelson1982,Winter2003}. To the
extent operating routines stems from idiosyncratic evolutionary process of individual
firms rather than being the result of investments in factor markets, they can satisfy the
VRIO condition and thus be a source of competitive advantage. In such instances, dynamic
capabilities work through operating routines in creating competitive advantage. This
indirect impact on competitive advantage is very much at the core of the DC literature
(ref). Firms with high dynamic capabilities will be able to better change operating
routines to facilitate strategic change in response to changing conditions facing the
firm. Dynamic capabilities then creates VRIO operating routines and hence competitive
advantage. When this happens, firms changes their operating routines in a evolutionary
making operating routines themselves OR itself satisfies the VRIO condition.

\emph{H2: Dynamic capabilities has an indirect effect on competitive advantage through
  changing the operating routines of the firm.}

On the other hand, resources and capabilities can to a certain extent be captured in
factor markets through investment decisions (ref) or through alliances (ref) or M\&A
activates (ref). By acquiring resources through investments, the operating routines of the
firms may not be directly affected. When these acquisitions happen as a response to
systematic resource orchestration (ref), i.e. reshaping the resource mix to obtain
coordinated resource deployments \citep{Kor2005,Pan2006}, they would indeed have potential
performance implications for the firm. Moreover, when this process of resource acquisition
process of DC develops a way to acquire and mix resource for the purpose of adapting to
change, the resources themselves would not necessarily satisfy the VRIO condition. The way
in which this happens may be much harder to imitate due to the complexities such routines
exhibit. Although resources themselves exhibit clear equifinality in their impact on
competitive advantage, combined with the simple routines of the orchestration
(i.e. dynamic capabilities), the sum would likely satisfy the VRIO condition. This is
especially true if routines are deployed rapidly in changing market conditions. In other
words, there are several ways in which resources can be orchestrated to generate CA,
whereas improved OR will always mean improved competitiveness et ceteris paribus. In such
instances DC itself can lead to CA directly.

\emph{H3: Dynamic capabilities has a direct effect on competitive advantage through the
  process of resource orchestration and acquisition of resources in factor markets.}

In other words, we would expect DC to have both a direct and an indirect impact on CA. The
indirect impact stems from the \emph{evolutionary path} of complex processes, whereas the
direct impact stems from the \emph{market path}.  Although both the direct and the
indirect path stems from DC creating, extending and modifying the resource base (OR or
factor market resources), the difference in the path’s stems from to what extent the
mediating construct alone can by itself satisfy the VRIO condition. Operating routines
can, due to their evolutionary nature, in turn shaped by DC, achieve CA
\citep{Collis1994,Winter2003}. Factor market resources and market acquired capabilities
and best practices can not \citep{Eisenhardt2000,Peteraf2013}, but in conjuncture with the
routines of acting in factor markets as well as the speed of deployment, the sum of
dynamic capabilities and factor market action would lead to competitive advantage.

However, the interlinkages between simple routines (the market path) and the coplex
processes (the evolutionary path) would vary with the environment.

\subsection{Interplay under environmental dynamism}

\cite{DiStefano2014} note that "even if specific simple rules are unstable and ephemeral,
the system as a whole is not" and that the "real source of sustainable competitive ad- vantage for an enterprise is the difficulty of imitating and substituting for the entire dynamic bundle that the system represents" (p.sss).

But the interlinkages between the parts of the system (the evolutionary- and the market
path), are likely to vary with the environmental dynamism. Under moderatly dynamism
dynamic capabilities takes the form of best pracitces (e.g. aquisition targeting) and is
less important than under more dynamic environments. However, albeit best practices, the
way they are implemented in the factor markets are still idiosyncratic. And taken into
conjunction with the routinzed process of changing underlying routines, VRIO will very
likely emerge under more stable conditions. It is not that the market/direct expression
of dynamic capabilities are not present in stable environments, rather it is less pronounced and the indirect
mediated effect (evolutionary path) is more important. In other words, it is not neither
nor, but degrees of importance.

Moreover, in high velocity environments the role of the evolutionary path tends to
weaken as such path-dependent changes occur slower compared to the market path. From the
market path with its simple routines we can argue that creation of "new knowledge” to
“allow for emergent adaptation” (EM, p. 1116) is the consequence. Consequently, the market
path is stronger and more prevalent under higher levels of environmental dynamism.

\emph{H4: Dynamic capabilities works relatively stronger through the market path in highly
dynamic environments.}

\begin{figure}
  \centering
  \captionsetup{width=0.5\linewidth}
  \includegraphics[width=0.5\linewidth]{./figure/fig2.png}
  \captionof{figure}{Dynamic capabilities direct and indirect paths to competitive advantage determined by the VRIO condition}
  \label{fig:fig2}
\end{figure}

Figure \ref{fig:fig2} illustrates the two proposed paths through which DC affect
competitive advantage. The main take away from this is that dynamic capabilities can
affect CA through changing operational routines that by themselves can satisfy the VRIO
condition by means of their evolutionary nature. Thus, we hypothesize that operating
routines mediates the relationship between dynamic capabilities and competitive advantage
the evolutionary path. Conversely, the same dynamic capabilities enable the firm to
identify resources and capabilities in factor markets that are less idiosyncratic and
often takes the shape of best practices that are definitely imitable and thus bot VRIO. In
such instances, the VRIO condition is satisfied by the DC itself through its complex
process of orchestrating and mixing resources – the market path.

\section{Data and methods}

To test the relationships between dynamic capabilities and competitive advantage we need a
proper metric of both. We also benefit vastly from studying the same firms over time.  The
lack of h longitudinal studies to investigate dynamics over time and control for path
dependency of organizational routines has been noted as permanent limitation for moving
the DC literature forward empirically and theoretically \cite{Schilke2018}.

We collected survey data from R\&D active firms in Norway at two points in time
($T_1=2005$ and $T_2=2014$). The population was all businesses registered to a scheme for
tax deduction of R\&D costs (called SkatteFUNN). As all enterprises which are eligible for
taxation could register their R\&D activities to receive a tax refund, the registered
enterprises include close to all enterprises which are involved in such activities at the
time of our study.

All firms that registered R\&D activities during May to December 2005 were approached
totaling in all 1721 firms. A web-based questionnaire was developed containing the
measures of the key constructs of this paper. A link to the questionnaire was e-mailed to
the firms within a month after they registered R\&D activities. The initial mailing was
followed by two e-mail reminders. Of the firms approached, 1199 (70 \%) returned filled-in
questionnaires. The 1199 companies that filled out the questionnaire were contacted again
9 years later. The majority of firms were contacted in spring/summer 2014. All received a
web-based questionnaire containing the same measures of the focal constructs of this
paper. 283 of the firms returned filled-in questionnaire.

All the firms in the sample was identifiable by means of the official firm
identifier. This enabled us to attach financial data from the annual accounts of the firms
despite none of the firms in the sample being publicly traded companies. We obtained the
financial accounts from the National Firm Registry (BRREG) for all years between 2007 and
2014 (prior to 2007 was not accessible through our database).

\subsection{Variable construction}

The focal constructs of this paper are dynamic capabilities, operational routines and
competitive advantage. These are all operationalized using the survey data described
above. Table 1 below describes the items used to operationalize the focal constructs and
their internal validity in forming an additive index (Cronbach´s alpha). The results show
that our key constructs can be represented by the items in the survey.

There are no established measurement models of dynamic capabilities
\citep{McKelvie2009,Schilke2018}. Therefore this study builds on a mixture of qualitative
case study methodology, literature review and statistical techniques to develop and refine
measures of DCs. First, exploratory qualitative interviews, using a semistructured
interview guide, were conducted with management representatives from 10 R\&D/innovative
firms. The aim of the interviews was to get an overview of each firm's innovation and
development processes, in particular the processes related to its dynamic capabilities and
the management of its resources. Themes raised in the interviews were about network,
co-operation with external R\&D-institutions, learning in the firm, adaptation and changes
in the firm. We interviewed SMEs and larger firms, and the industries varied from
high-tech and ICT to publishing.  Based on the interviews and an extensive literature
review, statements identified as descriptions of dynamic capabilities and resources were
developed and included in a questionnaire.

Second, the informants from the 10 firms were subsequently asked to take part in a pretest
of the questionnaire, including the preliminary items developed to measure dynamic
capabilities and resources, by responding to the questionnaire and giving comments on the
individual questions. This was followed up by a telephone call to the interviewees where
they were asked to report their views on the various questions/items.

Third, the face validity of the items was further examined by pre-testing the measurements
among experts. Researchers with knowledge of business strategy within firms were asked to
evaluate the questionnaire. Based on the results from the pilot study, the items were
adjusted and refined.

Although we recognize at the outset that the concept of dynamic capabilities and their
underlying resource components are very challenging to research in a systematic and
econometric fashion \citep{McKelvie2009}, we follow the argument in the
literature that more empirical work is necessary to test and refine the dynamic
capabilities concept and how it is related to the evolutionary economic theory
\citep{Arend2009,McKelvie2009} \todo{Rosenblom, 2000; Verona and Ravasi, 2003;}.  It
is in this spirit that the research reported in this paper has been undertaken.

Items measuring DC and OCs were developed as statements for which the respondents were
asked to indicate to what extent each statement fitted a description of their business. We
adopted one-sided seven point Likert scale where: 1 = strongly disagree and 7 = strongly
agree. We built on prior studies where items measuring OCs have been measured relative to
competitors \citep{McKelvie2009}.

<<result1, results="asis", echo=FALSE >>=

ht

@

\subsection{Control variables}

We add a set of control variables expected to explain parts of the variation in our focal
constructs. From the survey we add firm size, firm age, the level of dynamism facing the
firm. Size is the reported number of employees in the firm at $T_2$, and age is the number of
years since establishment. Dynamism is captured by asking a number of questions about the
competitive environment facing the firm. A full list of items and their internal validity
(0.75) is available in appendix 1111.

To partly remedy the problem of common method bias, we add controls of financial data from
the firm´s profit and loss statements, and balance sheet. Specifically, we add the natural
logarithm of the assets and debt of the firm to capture the leverage (debt risk) of the
firm, as well as the capital intensity.

One core critique of the dynamic capabilities literature is that it simply is a response
to the firm being a good firm \citep{Arend2009}. This tautology makes it hard to
distinguish the direction of the effects the performance implications of dynamic
capabilities. Simply put both firm performance and dynamic capabilities can be due to some
unobserved factor of the firm simply being a “good firm”. In such cases we would simply
regress one on the other and get positive results without the exact mechanism being
pointed out. To partly control for this we add a proxy for the quality of the firm as a
control variable. In this paper we use the profit margin of the firm as this type of
control. It is not perfect as profit margins can be due to many factors outside our model,
but it serves to control for parts of the possible cofounding factors of firm quality.

\subsection{Estimation and empirical model}

To test our hypotheses, we run a sequential linear regression with heteroskedastic robust
standard errors. We test the mediation hypotheses within this framework using a causal
mediation \citep{Baron1986,MacKinnon2008} with quasi-Bayesian confidence intervals and
over a million simulations \citep{Imai2010}.

In addition we estimate mediating models and test formally for mediation using the Baron
\& Kenny method \citep{Baron1986,MacKinnon2008}. We estimate two mediating model to test
if OR mediates the effect of DC on CA, and if DC mediates the effect of DL on OR,
respectively.

\section{Results}
The focal variables of our study is presented with pairwise correlation, mean and standard
deviation in table \ref{tab:desc}.It is worth noting that the mean value of OR and DC is
declining over time suggesting that the average firm experience a negative development in
core resources. The ability of the firms in our sample to generate CA is somewhat
unchanged over the sample period. We also observe a certain correlation in the development
in DC over time suggesting that prior period DC is a strong predictor of future
DC. Furthermore, we see a strong correlation between DL and DC which indicates that these
two constructs are indeed related.

\begin{figure}[ht] \label{ fig7} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{full.pdf} 
    \caption{Full model} 
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{ma.pdf} 
    \caption{Routine mediation} 
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{or.pdf} 
    \caption{Process mediation} 
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.5\linewidth}
    \includegraphics[width=.99\linewidth]{dir.pdf} 
    \caption{Direct model} 
  \end{minipage} 
\end{figure}

\begin{figure}
     \centering
  \caption{Structural model of full estimation - PLS-SEM}
  \input{./figure/fullmod.tex}
\end{figure}

\begin{table}
\begin{center}
\script size
\begin{tabular}{lcccc}

\hline \\[-2pt]
 & \multicolumn{4}{c}{{\bf Models:}} \\  \cline{2-5} \\[-3pt] 
 & \emph{Direct} & \emph{Process} & \emph{Routines} & \emph{Full}\\
 & (1) & (2) & (3) & (4)\\
\hline
 &  &  &  & \\
\(CA_{t-1} \rightarrow CA\) & \Sexpr{plmest[1,2]} & \Sexpr{plmest[1,3]} & \Sexpr{plmest[1,4]} & \Sexpr{plmest[1,5]}\\
 & (\Sexpr{plmt[1,2]}) & (\Sexpr{plmt[1,3]}) & (\Sexpr{plmt[1,4]}) & (\Sexpr{plmt[1,5]})\\[3pt]
 &  &  &  & \\
\(DC_P \rightarrow CA\) & \Sexpr{plmest[2,2]} & \Sexpr{plmest[2,3]} & \Sexpr{plmest[2,4]} & \Sexpr{plmest[2,5]}\\
 & (\Sexpr{plmt[2,2]}) & (\Sexpr{plmt[2,3]}) & (\Sexpr{plmt[2,4]}) & (\Sexpr{plmt[2,5]})\\[4pt]
\(DC_X \rightarrow CA\) & \Sexpr{plmest[3,2]} & \Sexpr{plmest[3,3]} & \Sexpr{plmest[3,4]} & \Sexpr{plmest[3,5]}\\
 & (\Sexpr{plmt[3,2]}) & (\Sexpr{plmt[3,3]}) & (\Sexpr{plmt[3,4]}) & (\Sexpr{plmt[3,5]})\\[4pt]
 &  &  &  & \\
\(DC_P \rightarrow P\) &  & \Sexpr{plmest[4,3]} &  & \Sexpr{plmest[4,5]}\\
 &  & (\Sexpr{plmt[4,3]}) &  & (\Sexpr{plmt[4,5]})\\[4pt]
\(DC_P \rightarrow R\) &  &  & \Sexpr{plmest[5,4]} & \Sexpr{plmest[5,5]}\\
 &  &  & (\Sexpr{plmt[5,4]}) & (\Sexpr{plmt[5,5]})\\[4pt]
\(DC_X \rightarrow P\) &  & \Sexpr{plmest[6,3]} &  & \Sexpr{plmest[6,5]}\\
 &  & (\Sexpr{plmt[6,3]}) &  & (\Sexpr{plmt[6,5]})\\[4pt]
\(DC_X \rightarrow R\) &  &  & \Sexpr{plmest[7,4]} & \Sexpr{plmest[7,5]}\\
 &  &  & (\Sexpr{plmt[7,4]}) & (\Sexpr{plmt[7,5]})\\[4pt]
 &  &  &  & \\
\(P \rightarrow CA\) &  & \Sexpr{plmest[8,3]} &  & \Sexpr{plmest[8,5]}\\
 &  & (\Sexpr{plmt[8,3]}) &  & (\Sexpr{plmt[8,5]})\\[4pt]
\(R \rightarrow CA\) &  &  & \Sexpr{plmest[9,4]} & \Sexpr{plmest[9,5]}\\
 &  &  & (\Sexpr{plmt[9,4]}) & (\Sexpr{plmt[9,5]})\\[4pt]
 &  &  &  & \\
\hline \\[-3pt] 
Goodness of fit & \Sexpr{gofdir} & \Sexpr{gofor} & \Sexpr{gofma} & \Sexpr{goffull}\\
Number of observations & \Sexpr{N} & \Sexpr{N} & \Sexpr{N} & \Sexpr{N}\\
  Firm level controls & YES & YES & YES & YES\\[3pt]
\hline
 &  &  &  & \\
\end{tabular}
\end{center}
 \label{tab:reg}
\end{table}

Table \ref{tab:reg} shows a summary of the regression models proposed in the previous
chapter. Models 1,2 and 3 are designed to test DL as \emph{antecedent} for $OR$ and
$DC$. Consider model 1 we observe that $DL$ is influencing the level of $OR$ within the
firm ($\beta_2 =$r) supporting H1. However, the effect disappears when we
add $DC$ in model 2 suggesting a mediating effect. In other words we see that firms with
initial high $DC$($\beta_3=$ \ {mod2_dc0}) and increasing $DC$ over time ($\beta_4=$
\ {mod2_ddc}) also have better $OR$. To control for the mediating effect suggested by
the change in coefficients for $DL$ between models 1 and 2, we performed a causal
mediation analysis as reported in figure \ref{fig:fig3}. This analysis confirms the
indications from models 1 and 2 showing that the direct effect from $DL$ on $OR$ is
insignificant and small (ADE = \ {medmod1_ade}). Meanwhile, the indirect effect
running through changes in $DC$ over time, is significant yielding a positive and
significant indirect effect of \ {medmod1_acme}. The proportion of the effect that is
mediated amounts to \ {medmod1_shmed}. In sum these results suggest that H1 is
unsupported, and that the bulk of the effect is indirect running through $DC$, in line with
the expectations in H3. Similarly, we see from model 3 that $DC$ is indeed influenced by $DL$
above and beyond its own evolutionary trajectory. This is in line with our expectations
and lend support to H4.

To evaluate H5 proposing a feedback effect from $DC$ to $DL$, we turn to model 6 in table
\ref{tab:reg}. Controlled for the usual firm level controls we find a positive effect
($\xi_1=$ \ {mod6_dc0}) from $DC_{T_1}$ on $DL$ measured at $T_1$. This supports H5
suggesting that firms with more $DC$ is more likely to generate opportunities where it
becomes more prevalent that learning will need to take place. One interesting finding is
also that this effect environmental dynamism decreases firms level of $DL$. This is indeed
puzzling and we will look in to this in a post-hoc analysis.\todo{When adding interactions
  we see that high environmental dynamism in conjunction with DC is positive for DL
  meaning that DC will indeed work to make firms more aware of the benefits of DL when
  environments are changing} It is also worth noting that models 1,2 and 3 have positive
and significant $\rho$ from the Heckman selection model indicating that selection is
leading to biased estimates. By running a Heckman correction we avoid this potential
problem.

Models 4 and 5 in table \ref{tab:reg} deals with the \emph{outcome} of DC. H6a and H6b
both considers effects of $DC$ on the competitive advantage of the firm $CA$. $CA$ is
influenced by the level of $DC$ of the firm indicating that dynamic capabilities are a
source of competitive advantage in line with our expectations. Furthermore, considering
model 5 we see that $OR$ also has a significant positive effect on $CA$ indicating that
operating routines can also be a source of advantage for the firm. Moreover, the
coefficient of $\Delta DC$ decline from  {mod4_ddc} to  {mod5_ddc} when we add
$OR$. This suggests that some mediation is at work, i.e. that changes in $DC$ leads to
changes in changes in $OR$ leading to $CA$. This indirect path is in line with the
expectations from our hypotheses and as described in figure \ref{fig:fig2}. When formally
testing this mediation running a causal mediation analysis (see figure \ref{fig:fig4}), we
find added support that both a \emph{direct-} and an \emph{indirect} path exist to explain
the relationship between $DC$ and $CA$ (ADE = \ {medmod5_ade}, ACME =
\ {medmod5_acme}) and a total mediated share of \ {medmod5_shmed}.

Table \ref{tab:hyptab} depict a summary of the hypotheses and indicate if they are
supported or not in our analysis. We will now turn to the discussion.

\section{Discussion}

Moving from antecedents to outcomes we extended the \cite{Zollo2002a} model by adding
competitive advantage to the model. We find that both operational capabilities and dynamic
capabilities can lead to competitive advantage. The most obvious path from the mere
definition of dynamic capabilities, runs through the evolution of operational
capabilities. Dynamic capabilities shape operational capabilities and changes them to
adapt, hence safeguarding that they are adapted. This is beyond the ad-hoc learning and
adaptation taking place as repetitive actions by employees. TO BE CONTINUED

\subsection{Limitations}

\section{Bibliography}

\singlespacing
\bibliography{Dissertation_clean}

\section{Appendix}

\end{document}