### 2019-11-18 figure/fullmod.tex

\begin{tikzpicture}
   \centering
   \tikzstyle{mynode}=[circle,draw, ultra thin, draw=black, fill=white, minimum
   size=17mm,inner sep=0pt, align=center]

   \tikzstyle{mytext}=[fill=none, minimum
   size=17mm,inner sep=0pt, align=center]

   \node[mynode] (dc1){$DC$};

   \node[mynode,right of=dc1, yshift= 2cm, xshift=2.5cm] (cap){$CAP$};
   \node[mynode,right of=dc1, yshift=-2cm,xshift=2.5cm] (cog){$COG$};
   \node[mynode,right of=dc1, xshift=6cm] (ca){$CA$};
   \node[mytext,above of=dc1,rotate=36, xshift=1.6cm,yshift=0cm] (ro){\bf Conductor};
   \node[mytext,above of=dc1,rotate=-36, xshift=2.8cm,yshift=-1.5cm] (cp){\bf Coach};

   \draw[-latex] (dc1.north east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
   fill=none] {} (cap.west);

    \draw[-latex] (dc1.south east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
     fill=none] {} (cog.west);

    \draw[-latex] (cap.east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
      fill=none] {} (ca.north west);

    \draw[-latex] (cog.east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
    fill=none] {} (ca.south west);

        \draw[-latex] (dc1.east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
      fill=none] {} (ca.west);

   %\draw[-latex] (dc2.east) -- node[text width=0.5cm,font=\footnotesize, above=2pt, align=left,
   %fill=none] {} (cap.west);

   %\draw[-latex] (dc1.south east) -- node[text width=0.5cm,font=\footnotesize,  above=2pt,
   %align=left, fill=white,sloped, near end] {$0.496^{***}$} (ma.north west);

   %\draw[-latex] (dc2.north east) -- node[text width=0.5cm,font=\footnotesize, above=2pt,align=left,
   %fill=white,sloped, near end]{$0.343^{***}$} (or.south west);

    %\draw[-latex] (or.east) -- node[text width=0.5cm,font=\footnotesize, align=left,above=2pt,
   %fill=white,sloped]{$0.185^{***}$} (ca.north west);

   %\draw[-latex] (ma.east) -- node[text width=0.5cm,font=\footnotesize, below=2pt,
   %fill=white,sloped]{$0.123^{**}$} (ca.south west);

   %\draw[-latex] (dc1.north)[bend left] to node[text width=0.5cm,font=\footnotesize, above=2pt,align=center,

   %fill=white]{$0.299^{***}$}(ca.north);

   %\draw[-latex] (dc2.south)[bend right] to node[text width=0.5cm,font=\footnotesize,below=2pt, align=left,
   %fill=white]{$-0.066^{}$} (ca.south);

\end{tikzpicture}