### 2020-02-25 figure/cortab.rnw
submitted dissertation

\singlespacing
\tiny
\setlength{\tabcolsep}{3pt}
<<corcalc, results="asis", echo=FALSE>>=

cormat <- mainf$cormat
rownames(cormat) <- c("1) $CA_1$",
                     "2) $CA_0$",
                     "3) $DC_0$",
                     "4) $CAP_1$",
                     "5) $CAP_0$",
                     "6) $COG_1$",
                     "7) $COG_0$",
                     "8) Size",
                     "9) ln(Assets)",
                     "11) Margin",
                     "12) Age")

names(cormat) <- c("Mean","SD",paste(seq(1:(length(cormat)-2)),sep=","))

print(xtable(cormat ,size="\\small",caption="Correlation matrix \\label{tab:cor}", align = c("l", rep("p{0.290in}",(dim(cormat)[2]))),table.placement="h", size = "\\small", tabular.environment="longtable"), size="scriptsize", label='tab:cor', caption.placement="top", floating=FALSE,latex.environments=NULL,sanitize.text.function=function(x){x})
@